|

### 2022-05-28 article/article.Rmd
pesky typos fixed

In this paper, we propose a theory based on executive leadership and party
politics to explain how city council size impacts local public goods provision.
We incorporate partisanship and mayoral first-move advantage to a
\citet{baron1989bargaining} bargaining model. Mayors make the initial policy
proposal, which consists of investments in public goods provision, rents for
personal gains, and a vector of side transfers for councilors. If mayors fail
to secure majoritarian support from councilors, a reversal stage occurs, in
which the decision leaves the mayor's office and councilors propose the public
investments. In this bargaining process, costs come from the side transfers
required to secure a majority among councilors. We show that in a model without
partisan concerns, bargaining costs always increase with larger city councils,
which in turn lower public goods provision and rents. However, partisanship
decreases bargaining costs when the chances of electing government-aligned
councilors are sufficiently high. Lower bargaining costs release funds for
mayors to invest in public goods. In this case, a larger city council may add
more members to the government coalition, thus reducing the mayor's bargaining
costs and increasing the equilibrium public goods provision.

@@

This paper contributes to four strands of the literature. First, this paper
dialogues with a large literature on multilateral bargaining. Repeated models
of multilateral bargaining started with \citet{baron1989bargaining}, and
scholars have recently analyzed how endogenous recognition power, the size of
the required majority, and partisanship impact these results
\citep{agranov2020persistence, baranski2020determinants,
calvert2005legislative, choate2019partisan}. Here we include the mayor's first
move advantage and political partisanship to Baron and Ferejohn
(\citeyear{baron1989bargaining}) and investigate the often-overlooked effect of
legislature size on bargaining equilibrium strategies.

>>>>

In this paper, we propose a theory based on executive leadership and party
politics to explain how city council size impacts local public goods provision.
We incorporate partisanship and mayoral first-mover advantage to
Baron and Ferejohn's (\citeyear{baron1989bargaining}) bargaining
model. Mayors make the initial policy proposal, which consists of investments
in public goods provision, rents for personal gains, and a vector of side
transfers for councilors. If mayors fail to secure majoritarian support from
councilors, a reversal stage occurs, in which the decision leaves the mayor's
office and councilors propose the public investments. In this bargaining
process, costs come from the side transfers required to secure a majority among
councilors. We show that in a model without partisan concerns, bargaining costs
always increase with larger city councils, which in turn lower public goods
provision and rents. However, partisanship decreases bargaining costs when the
chances of electing government-aligned councilors are sufficiently high. Lower
bargaining costs release funds for mayors to invest in public goods. In this
case, a larger city council may add more members to the government coalition,
thus reducing the mayor's bargaining costs and increasing the equilibrium
public goods provision.

@@

This paper contributes to four strands of the literature. First, this paper
dialogues with a large literature on multilateral bargaining. Repeated models
of multilateral bargaining started with \citet{baron1989bargaining}, and
scholars have recently analyzed how endogenous recognition power, the size of
the required majority, and partisanship impact these results
\citep{agranov2020persistence, baranski2020determinants,
calvert2005legislative, choate2019partisan}. Here we include the mayor's
first-mover advantage and political partisanship to Baron and Ferejohn's
(\citeyear{baron1989bargaining}) model and investigate the often-overlooked
effect of legislature size on bargaining equilibrium strategies.
